---
title: "Taxon of Interest"
author: "Yaqian Luo"
date: "March 9, 2018"
output: html_document
---
***Methods***
```{r}
library(tidyverse)
library(phyloseq)
library(magrittr)
```

Based on your initial exploration of the data, you will choose 1 taxon to investigate in-depth. You can choose any taxon from any taxonomic level under the following criteria:
*Not an "unclassified" group
*Present at abundance > 0 in a least 3 samples
*Encompasses at least 5 OTUs/ASVs ???You may also want to give yourself an upper limit on number of OTUs/ASVs as too many will make your plots large and unwieldy, though we are not specifically setting a maximum for you.

Abundance for each phylum in seven samples from different depths.

***Mothur***

```{r}
library(phyloseq)
library(tidyverse)
load("mothur_phyloseq.RData")
mothur
mothur_percent = transform_sample_counts(mothur, function(x) 100 * x/sum(x))
plot_bar(mothur_percent, fill="Phylum") +
  geom_bar(aes(fill=Phylum), stat="identity") +
  facet_wrap(~Phylum, scales="free_y") +
  theme(legend.position="none")
```

***QIIME2***
```{r}
library(tidyverse)
load("qiime2_phyloseq.RData")
qiime2
qiime2_percent = transform_sample_counts(qiime2, function(x) 100 * x/sum(x))
plot_bar(qiime2_percent, fill="Phylum") +
  geom_bar(aes(fill=Phylum), stat="identity") +
  facet_wrap(~Phylum, scales="free_y") +
  theme(legend.position="none")
```

***Data cleaning***
```{r}
# mothur data
set.seed(4832)
m.norm = rarefy_even_depth(mothur, sample.size=100000)
m.perc = transform_sample_counts(m.norm, function(x) 100 * x/sum(x))

# qiime2 data
set.seed(4832)
q.norm = rarefy_even_depth(qiime2, sample.size=100000)
q.perc = transform_sample_counts(q.norm, function(x) 100 * x/sum(x))
```

Question 1: How does microbial community structure change with depth and oxygen concentration?

***Alpha diversity***
mothur data
```{r}
m.alpha = estimate_richness(m.norm, measures = c("Chao1", "Shannon"))
m.meta.alpha = full_join(rownames_to_column(m.alpha), rownames_to_column(data.frame(m.perc@sam_data)), by = "rowname")

# Alpha-diversity across depth
m.meta.alpha %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=Shannon)) +
   geom_smooth(method='auto', aes(x=as.numeric(Depth_m), y=Shannon)) +
  labs(title="Figure 1: Alpha-diversity across depth using mothur data", y="Shannon's diversity index", x="Depth (m)")

# Alpha-diversity across oxygen
m.meta.alpha %>% 

ggplot() +
  geom_point(aes(x=O2_uM, y=Shannon)) +
  labs(title="Figure 2: Alpha-diversity across oxygen using mothur data", y="Shannon's diversity index", x="Oxygen (uM)")

# Alpha-diversity by oxic/anoxic
m.meta.alpha %>% 
  mutate(O2_group = ifelse(O2_uM == 0, "anoxic", "oxic")) %>% 

ggplot() +
  geom_boxplot(aes(x=O2_group, y=Shannon)) +
  labs(title="Figure 3: Alpha-diversity by oxic/anoxic using mothur data", y="Shannon's diversity index", x="Oxygen")
```

Qiime2 data
```{r}
q.alpha = estimate_richness(q.norm, measures = c("Chao1", "Shannon"))
q.meta.alpha = full_join(rownames_to_column(q.alpha), rownames_to_column(data.frame(q.perc@sam_data)), by = "rowname")

# Alpha-diversity across depth
q.meta.alpha %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=Shannon)) +
   geom_smooth(method='auto', aes(x=as.numeric(Depth_m), y=Shannon)) +
  labs(title="Figure 4: Alpha-diversity across depth using qiime2 data", y="Shannon's diversity index", x="Depth (m)")

# Alpha-diversity across oxygen
q.meta.alpha %>% 

ggplot() +
  geom_point(aes(x=O2_uM, y=Shannon)) +
  labs(title="Figure 5: Alpha-diversity across oxygen using qiime2 data", y="Shannon's diversity index", x="Oxygen (uM)")

# Alpha-diversity by oxic/anoxic
q.meta.alpha %>% 
  mutate(O2_group = ifelse(O2_uM == 0, "anoxic", "oxic")) %>% 

ggplot() +
  geom_boxplot(aes(x=O2_group, y=Shannon)) +
  labs(title="Figure 6: Alpha-diversity by oxic/anoxic using qiime2 data", y="Shannon's diversity index", x="Oxygen")
```

***Taxa presence and abundance***

Mothur data
```{r}
# Domain across samples
m.perc %>% 
  
plot_bar(fill="Domain") + 
  geom_bar(aes(fill=Domain), stat="identity") +
  labs(title="Figure 7: Domains across samples using mothur data")

#Phyla across samples
m.perc %>% 
  
plot_bar() + 
  geom_bar(aes(fill=Domain), stat="identity") +
  facet_wrap(~Phylum, scales="free_y")+
  labs(title="Figure 8: Phyla across samples using mothur data")

# Domain boxplots
m.perc %>%
  tax_glom(taxrank = 'Domain') %>%
  psmelt() %>% 

ggplot() +
  geom_boxplot(aes(x=Domain, y=Abundance)) +
  coord_flip() +
  labs(title="Figure 9: Domain boxplots using mothur data")

```
qiime2 data
```{r}
# Domain across samples
q.perc %>% 
  
plot_bar(fill="Domain") + 
  geom_bar(aes(fill=Domain), stat="identity") +
  labs(title="Figure 10: Domains across samples using qiime2 data")

#Phyla across samples
q.perc %>% 
  
plot_bar() + 
  geom_bar(aes(fill=Domain), stat="identity") +
  facet_wrap(~Phylum, scales="free_y")+
  labs(title="Figure 11: Phyla across samples using qiime2 data")

# Domain boxplots
q.perc %>%
  tax_glom(taxrank = 'Domain') %>%
  psmelt() %>% 

ggplot() +
  geom_boxplot(aes(x=Domain, y=Abundance)) +
  coord_flip() +
  labs(title="Figure 12: Domain boxplots using qiime2 data")
```

Question 3: Within your taxon, what is the richness (number of OTUs/ASVs)?

Mothur data
```{r}
m.norm %>% 
  subset_taxa(Phylum=="Cyanobacteria")%>%
  estimate_richness(measures = c("Observed"))
```

qiime data
```{r}
q.norm %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria")%>%
  estimate_richness(measures = c("Observed"))
```


***Numbers of OTUs/ASVs***
```{r}
library(kableExtra)
library(knitr)
number = data.frame(
  number = c(1,2,3,4,5,6,7,8,9),
  Phylum = c("Bacteroidetes", "Chlamydiae", "Chloroflexi","Cyanobacteria","Firmicutes","Gemmatimonadetes","Thaumarchaeota","Verrucomicrobia","Proteobacteria"),
  OTU = c(503,13,46,22,19,17,94,82,2101),
  ASV = c(269,18,47,52,22,7,27,67,1121)
)

number %>% 
  kable("html") %>%
  kable_styling(bootstrap_options = "striped", font_size = 10, full_width = F)
```

***Results***
1.How does microbial community structure change with depth and oxygen concentration? 

*Alpha-diversity

*a)Simpson Reciprocal Index (SRI) using mothur data*

```{r}
library(vegan)
library(dplyr)
library(ggplot2)
# Calculate SRI
SRI_mothur=data.frame(diversity(OTU_mothur, index="invsimpson")) 
names(SRI_mothur)=c("SRI")

# Table of Simpson Reciprocal Index (SRI) with depth and oxygen using mothur data
SRI_mothur_depth_oxygen=cbind(SRI_mothur,sam_mothur)%>% select("SRI","Depth_m","O2_uM")
SRI_mothur_depth_oxygen

# Figure of Simpson Reciprocal Index (SRI) with depth using mothur data
SRI_mothur_depth_oxygen%>%
  ggplot(aes(x=Depth_m, y=SRI))+
  geom_point()

# Figure of Simpson Reciprocal Index (SRI) with oxygen using mothur data
SRI_mothur_depth_oxygen%>%
  ggplot(aes(x=O2_uM, y=SRI))+
  geom_point()
```

 *b)Simpson Reciprocal Index (SRI) using qiime data*
```{r}
library(vegan)
library(dplyr)
library(ggplot2)
# Calculate SRI
OTU_qiime2_t=t(OTU_qiime2)

SRI_qiime2=data.frame(diversity(OTU_qiime2_t, index="invsimpson")) 
names(SRI_qiime2)=c("SRI")

# Table of Simpson Reciprocal Index (SRI) with depth and O2 using qiime2 data
SRI_qiime2_depth_O2=cbind(SRI_qiime2,sam_qiime2)%>% select("SRI","Depth_m","O2_uM")
SRI_qiime2_depth_O2

# Figure of Simpson Reciprocal Index (SRI) with depth using mothur data
SRI_qiime2_depth_O2 %>%
  ggplot(aes(x=Depth_m, y=SRI))+
  geom_point()

# Figure of Simpson Reciprocal Index (SRI) with O2 using mothur data
SRI_qiime2_depth_O2 %>%
  ggplot(aes(x=O2_uM, y=SRI))+
  geom_point()
```

???Taxa presence and abundance

2.Does your taxon of interest significantly differ in abundance with depth and/or oxygen concentration?

3.Within your taxon, what is the richness (number of OTUs/ASVs)?
 **number of OTUs in cyanobateria**
```{r}
mothur
OTU_mothur=data.frame(mothur@otu_table)
tax_mothur=data.frame(mothur@tax_table)
sam_mothur=data.frame(mothur@sam_data)

library(dplyr)
OTU_richness=filter(tax_mothur,Phylum=="Cyanobacteria")%>%count()
  
```
**number of ASVs in cyanobateria**
```{r}
qiime2
OTU_qiime2=data.frame(qiime2@otu_table)
tax_qiime2=data.frame(qiime2@tax_table)
sam_qiime2=data.frame(qiime2@sam_data)

library(dplyr)
ASV_richness=filter(tax_qiime2,Phylum=="Cyanobacteria")%>%count()
```

4.Do the abundances of OTUs/ASVs within your taxon of interest change significantly with depth and/or oxygen concentration?

mothur data
```{r}
# Abundance of OTUs within cyanobacteria phylum across depth
m.perc %>% 
  subset_taxa(Phylum=="Cyanobacteria") %>% 
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=Abundance)) +
  geom_smooth(method='lm', aes(x=Depth_m, y=Abundance)) +
  facet_wrap(~OTU, scales="free_y") +
  labs(title="Figure 15: Abundance of OTUs within cyanobacteria phylum across depth")

m.perc %>% 
  subset_taxa(Phylum=="Cyanobacteria") %>%
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=OTU, size=Abundance, color=OTU)) + 
  scale_size_continuous(range = c(1,6)) +
  labs(title="Figure 16: Abundance of OTUs within cyanobacteria phylum across depth")

#Abundance of OTUs within cyanobacteria phylum across oxygen
m.perc %>% 
  subset_taxa(Phylum=="Cyanobacteria") %>% 
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=O2_uM, y=Abundance)) +
  geom_smooth(method='lm', aes(x=O2_uM, y=Abundance)) +
  facet_wrap(~OTU, scales="free_y") +
  labs(title="Figure 17: Abundance of OTUs within cyanobacteria phylum across oxygen")

m.perc %>% 
  subset_taxa(Phylum=="Cyanobacteria") %>%
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=O2_uM, y=OTU, size=Abundance, color=OTU)) + 
  scale_size_continuous(range = c(1,6)) +
  labs(title="Figure 18: Abundance of OTUs within cyanobacteria phylum across oxygen")

```

qiime2 data
```{r}
# Abundance of ASVs within cyanobacteria phylum across depth
q.perc %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>% 
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=Abundance)) +
  geom_smooth(method='lm', aes(x=Depth_m, y=Abundance)) +
  facet_wrap(~OTU, scales="free_y") +
  labs(title="Figure 19: Abundance of ASVs within cyanobacteria phylum across depth")

q.perc %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>%
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=OTU, size=Abundance, color=OTU)) + 
  scale_size_continuous(range = c(0,5)) +
  labs(title="Figure 20: Abundance of ASVs within cyanobacteria phylum across depth")

#Abundance of OTUs within cyanobacteria phylum across oxygen
q.perc %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>% 
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=O2_uM, y=Abundance)) +
  geom_smooth(method='lm', aes(x=O2_uM, y=Abundance)) +
  facet_wrap(~OTU, scales="free_y") +
  labs(title="Figure 21: Abundance of ASVs within cyanobacteria phylum across oxygen")

q.perc %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>%
  psmelt() %>% 
  
ggplot() +
  geom_point(aes(x=O2_uM, y=OTU, size=Abundance, color=OTU)) + 
  scale_size_continuous(range = c(0,5)) +
  labs(title="Figure 22: Abundance of ASVs within cyanobacteria phylum across oxygen")

```

Question 2: 

```{r}
library(dplyr)
library(ggplot2)
q.norm %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>% 
  tax_glom(taxrank = 'Phylum') %>%
  psmelt() %>%

  lm(Abundance ~ Depth_m, .) %>% 
  summary()

q.perc %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>% 
  psmelt() %>% 
  group_by(Sample) %>% 
  summarize(Abundance_sum=sum(Abundance), Depth_m=mean(Depth_m)) %>% 
  
ggplot() +
  geom_point(aes(x=Depth_m, y=Abundance_sum)) +
  geom_smooth(method='lm', aes(x=as.numeric(Depth_m), y=Abundance_sum)) +
  labs(title="Example 7: Abundance unclassified domain across depth using qiime2 data")
```

```{r}
library(dplyr)
library(ggplot2)
q.norm %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>% 
  tax_glom(taxrank = 'Phylum') %>%
  psmelt() %>%

  lm(Abundance ~ O2_uM, .) %>% 
  summary()

q.perc %>% 
  subset_taxa(Phylum=="D_1__Cyanobacteria") %>% 
  psmelt() %>% 
  group_by(Sample) %>% 
  summarize(Abundance_sum=sum(Abundance), O2_uM=mean(O2_uM)) %>% 
  
ggplot() +
  geom_point(aes(x=O2_uM, y=Abundance_sum)) +
  geom_smooth(method='lm', aes(x=as.numeric(O2_uM), y=Abundance_sum)) +
  labs(title="Example 7: Abundance unclassified domain across oxygen using qiime2 data")
```
```

5.Are the answers to the above the same using mothur and QIIME2 processed data?

