---
title: "Assignment_Plotting_data_in_R"
author: "Yaqian Luo (59751503)"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

## Exercise 1

Plot another nutrient of your choice against depth. Change the points to be purple triangles.

```{r}
library(ggplot2)
metadata=read.table(file="Saanich.metadata.txt", header=TRUE, row.names=1, sep="\t", na.strings="NAN")
ggplot(metadata, aes(x=PO4_uM,y=Depth_m))+
  geom_point(shape=17, color="purple")
```

## Exercise 2

Using dplyr, convert the temperature variable from Celsius to Fahrenheit. Then create a dot plot of temperature in Fahrenheit against depth.

```{r}
library(dplyr)
library(ggplot2)
metadata%>%
  mutate(Temperature_F=Temperature_C*1.8+32) %>%
  ggplot(aes(x=Temperature_F,y=Depth_m)) +
    geom_point()
```

## Exercise 3

Create a bar plot at a different taxonomic level with more descriptive x- and y-axis labels as well as a title. Use the ggplot cheatsheet to learn the guides needed to change these labels.

```{r}
library(phyloseq)
library(ggplot2)
load("phyloseq_object.RData")
physeq_percent=transform_sample_counts(physeq,function(x) 100*x/sum(x))
plot_bar(physeq_percent,fill="Domain")+
  geom_bar(aes(fill=Domain),stat="identity")+
  labs(x="Sample depth", y="Percent relative abundance", title="Domain from 10 to 200 m in Saanich Inlet")
```

## Exercise 4

Using ggplot, create a faceted figure showing nutrient concentrations in uM for O2, PO4, SiO2, NO3, NH4, and NO2 by depth. For this, you can use either the metadata table or the phyloseq object.

Hint: Explore the dplyr function gather to manipulate the metadata so that your nutrients of interest are all in one variable (column).

```{r}
library(tidyverse)
library(dplyr)
nutrient=metadata %>% select(Depth_m,O2_uM,PO4_uM,SiO2_uM,NO3_uM,NH4_uM,NO2_uM) %>% gather(nutrient,uM,O2_uM,PO4_uM,SiO2_uM,NO3_uM,NH4_uM,NO2_uM)
library(ggplot2)
ggplot(nutrient,aes(x=Depth_m,y=uM))+
  geom_line()+
  geom_point()+
  facet_wrap(~nutrient,scales="free_y")
```

